---
title: "Assignment 4"
author: "Emma Colley"
date: "10/07/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries, message = FALSE}
library(osmdata)
library(opentripplanner)
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(viridis)
library(dplyr)
```


# Introduction

I downloaded data points of all housing developments completed in 2020 that include at least 20 affordable units. In this exercise, I will investigate spatial relationships between these housing developments and stops along the T.


```{r, results='hide'}
Tstops <- st_read("C:\\Users\\emmac\\Documents\\GitHub\\emmacolley-vis\\T_stops\\MBTA_Rapid_Transit.shp")
```

```{r, results='hide'}

affordable_2020 <- st_read("C:\\Users\\emmac\\Documents\\GitHub\\emmacolley-vis\\affordable_2020\\massbuilds-shp-20201004-a44e92.shp") %>%
  filter(MUNICIPAL == "Boston") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") #project to wgs84


MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"
```



```{r, results='hide'}
opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/boston_streets.osm')
```

```{r, results='hide'}
path_otp <- otp_dl_jar("OTP")
```

```{r, results='hide'}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024)
```


```{r, results='hide'}
otp_setup(otp = path_otp, dir = path_data, memory = 1024)

# Connect to opentripplanner
otpcon <- otp_connect()
```
## 1) Creating Isochrones


I plotted the new housing developments with affordable units and created two isochrones: a five minute walk and a five minute drive. I've also layered in the T stops (all train lines) to begin to see how close they are to the housing developments.



```{r, results='hide'}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = affordable_2020,
                mode = "WALK", cutoffSec = 300) %>%
                st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = affordable_2020, 
                mode = "CAR", cutoffSec = 300) %>%
                st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)
```

```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = affordable_2020) +
  geom_sf(data = Tstops, color = "tomato1") +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_manual(values = c("skyblue", "royalblue"), name = "Area that is reachable\nwithin 5 min", labels = c("by car", "by walking")) +
  theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", .5))) +
  labs(caption = "Red dots represent T stops.\nBasemap Copyright OpenStreetMap contributors.")
```

## 2) Plotting Walking and Driving distances


I've labeled each housing development with their neighborhood to begin to visual any trends between neighborhoods and commuting distances. Which neighborhoods have larger walk sheds? Which have large drive sheds? It's hard to distinguish a pattern since I have a small sample set.



```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(drive))) +
  geom_point(color = "lightslategrey", size = 3) +
  geom_text(aes(label = c("Roxbury", "South Boston", "South Boston", "Dorchester", "Mission Hill", "Brighton", "Jamaica Plain","Jamaica Plain","South End","South End")), hjust=-.2, size = 3) +
  scale_x_continuous(name = "Area within a five-minute walking distance\nof housing development (square km)",
            breaks = breaks <- seq(10000, 200000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute driving distance\nof housing developmen (square km)",
            breaks = breaks <- seq(0, 700000, by = 100000),
            labels = breaks / 1000000) +
  theme_minimal()
```


## 3) Assigning a Transit Score to Each Isochrone


I will use now investigate how many T stops are within each isochrone to see which housing developments are more accessible to the T by walking or driving. I am including both isochrones in this study. 


```{r, results='hide', message=FALSE}
st_as_sf(iso_all_modes)
iso_transformed <- st_transform(iso_all_modes, crs = "+proj=longlat +datum=WGS84")
```

```{r, results='hide', message=FALSE}
iso_transformed <- iso_transformed %>%
  mutate(transit_score = lengths(st_covers(geometry, Tstops)))
```
```{r}
ggplot(iso_transformed) +
  annotation_map_tile(zoomin = 1, type = "cartolight", progress = "none") +
  geom_sf(aes(fill=transit_score), alpha=.5) +
  scale_fill_viridis_c(option = "C", name = "Walk and drive sheds\nwith T stops", breaks = breaks <- seq(0, 4, by = 1), labels = paste(prettyNum(breaks), "T stops")) +
  theme_map() 
```


Ok, that figure is a bit helpful but it looks like most of the development's walk and drive sheds only have 1 T stop within them. I will take a step back and compare isochrones with and without T stops.



## 4) Which Isochrones overlap with a T stop?


6 of the 10 isochrones have a T stop within them. That means it would take 5 minutes to walk or 5 minutes to drive to the T from 6 of these affordable housing developments. They are shaded in blue in the figure below.


```{r, result='hide'}
iso_transformed <- iso_transformed %>%
  mutate(num_Tstops = lengths(st_covers(iso_transformed, Tstops))) %>%
  mutate(has_Tstops = num_Tstops > 0)
```
```{r}
n_Tstops_iso_transformed <- sum(iso_transformed$has_Tstops)

n_Tstops_iso_transformed
```

```{r}
ggplot(iso_transformed) +
  annotation_map_tile(zoomin = 1, type = "cartolight", progress = "none") +
  geom_sf(data = iso_transformed,
          aes(fill = has_Tstops)) +
  scale_fill_manual(values = c("gray85", "darkblue"),
          name = "Boston Neighborhoods\nby presence of a community center", 
          labels = c("Without a T stop",
                     "With a T stop")) +
  theme_map()
```


```{r, results='hide'}
otp_stop()
```

