---
title: "Assignment 21"
author: "Emma Colley"
date: "9/16/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
---
# Introduction

I created a map of the conterminous United States with its associated subareas (states) and displayed this area using three different map projections. I examined a state-level variable of Eviction Rate, provided to me by The Eviction Lab, and used three different methods to show how it varies from state to state.

This project uses data from The Eviction Lab at Princeton University, a project directed by Matthew Desmond and designed by Ashley Gromis, Lavar Edmonds, James Hendrickson, Katie Krywokulski, Lillian Leung, and Adam Porton. The Eviction Lab is funded by the JPB, Gates, and Ford Foundations as well as the Chan Zuckerberg Initiative. More information is found at evictionlab.org.


## Loading Libraries

```{r load libraries, message = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
```

## Map of the conterminous United States

```{r}
US_states <- ne_states(country = "United States of America", 
                       returnclass = "sf") %>%
  filter(name != "Alaska",
         name != "Hawaii")

US_map <- ggplot(US_states) +
  geom_sf() +
  theme_bw()

US_map
```
```{r}
#United States
USA_AEA <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=clrk66 +units=m +no_defs"
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"
HI_AEA <- "+proj=aea +lat_1=8 +lat_2=18 +lat_0=13 +lon_0=-157 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
```

## Reprojecting based on USA Contiguous Albers Equal Area Conic

```{r}
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = USA_AEA) +
  theme_bw()
```

## Reprojecting based on MA State Plan

```{r}
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = MA_state_plane) +
  theme_bw()
```

## Reprojecting based on Hawaii Albers Equal Area Conic

```{r}
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = HI_AEA) +
  theme_bw()
```
```{r}
states_transformed <- US_states %>%
  st_transform(crs = USA_AEA)
```

```{r}
ggplot(states_transformed) +
  geom_sf() +
  theme_bw()
```
```{r, results = 'hide', message = FALSE}
evictions_us <- read_csv("evictions_states.csv") %>%
  filter(year == 2016)

evictions_states <-states_transformed %>%
  left_join(evictions_us, by = "name") %>%
  select(name, postal, population, `eviction-rate`, `pct-af-am`)
  
```

## Chloropleth Map

I am mapping the eviction rate of each state in the contiguous United States. The highest eviction rate is in South Carolina at 8.87% and the lowest in is New Jersey at 0.10%. My classmate Megan Willis-Jackson helped me label the states.


```{r, results = 'hide', message = FALSE}
ggplot(evictions_states, aes(fill = `eviction-rate`)) +
  geom_sf(color = NA) +
  labs(title = "2016 Eviction Rates in the United States") +
  labs(caption = "Note: North Dakota, South Dakota, and Arkansas are missing data in 2016.") +
  scale_fill_viridis_c(name = "Eviction Rate") +
  geom_sf_label(aes(geometry, label = postal), size = 1.8, color = "white", fill = NA, label.size = NA, fontface = "bold",)
  theme_map() +
  theme(legend.background = element_blank())
```

## Continuous Cartogram Sized by Total Population

Here I am still color coding the states by Eviction Rate but I am distorting the shapes of the states by population (total number of people.)


```{r, results = 'hide', message = FALSE}
evictions_cartogram_cont <- evictions_states %>%
  cartogram_cont("population")

ggplot(evictions_cartogram_cont, aes(fill = `eviction-rate`)) +
  geom_sf(color = NA) +
  labs(title = "2016 Eviction Rates in the United States, States Altered by Population Size") +
  labs(caption = "Note: North Dakota, South Dakota, and Arkansas are missing data in 2016.") +
  scale_fill_viridis_c(name = "Eviction Rate") +
  theme_map() +
  theme(legend.background = element_blank())
```

## Continuous Cartogram Sized by Percentage of Population that is African American

Here I am still color coding the states by Eviction Rate but I am distorting the shapes of the states by the percentage of the population that is African American. I was predicting that these two variables would be telling when paired, but this cartogram is pretty illegible. It does look like the "larger" states in the South and South East with a high percentage of African Americans have medium to high eviction rates. 


```{r, results = 'hide', message = FALSE}
evictions_cartogram_cont <- evictions_states %>%
  cartogram_cont("pct-af-am")

ggplot(evictions_cartogram_cont, aes(fill = `eviction-rate`)) +
  geom_sf(color = NA) +
    labs(title = "2016 Eviction Rates in the United States, States Altered by African American Population Size") +
  labs(caption = "Note: North Dakota, South Dakota, and Arkansas are missing data in 2016.") +
  scale_fill_viridis_c(name = "Eviction Rate") +
  geom_sf_label(aes(geometry, label = postal), size = 1.8, color = "white", fill = NA, label.size = NA, fontface = "bold",) +
  theme_map() +
  theme(legend.background = element_blank())
```

## Proportional Symbol Map

Here is a proportional dot map. It looks like there are high eviction rates along the east coast, especially in the South East, but the representation could be misleading because those states are also small and close together.

```{r, results = 'hide', message = FALSE}
evictions_centeroids <- evictions_states %>%
  st_centroid()

ggplot(states_transformed) +
  geom_sf(fill = NA, color = "gray") +
  labs(title = "2016 Eviction Rates in the United States") +
  labs(caption = "Note: North Dakota, South Dakota, and Arkansas are missing data in 2016.") +
  geom_sf(data = evictions_centeroids, 
          aes(size = `eviction-rate`),
          alpha = 0.5, color = "green") +
  scale_size_continuous(name = "Eviction Rate", range = c(0, 16)) +
  theme_void() 
```

